---
title: "HKU3-CoV"
author: "Ellen Risemberg"
date: "3/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages 

```{r message=FALSE, warning=FALSE}
library(knitr)
library(qtl)
library(MESS)
library(MASS)
library(cowplot)
library(ggplot2)
library(scales)
source('code-dependencies/qtl_functions.R')
```

ggplot stuff:

```{r}
load_themes()
```

## Load data 

```{r}
HKU3 <- read.cross(format="csv", file='derived_data/rqtl_files//HKU3_CC006xCC044_rqtl.csv', 
                      na.strings=c("-","NA", "not tested"), genotypes=c("AA","AB","BB"), 
                      alleles=c("A","B"))
HKU3 <- jittermap(HKU3)
summary(HKU3)
#plotMap(HKU3) 

HKU3$pheno$batch <- as.factor(HKU3$pheno$batch) # change batch to factor
```


## Calculate derived measures 

**Calculate percentage of starting weight**

```{r}
HKU3$pheno$pd0 <- 100
HKU3$pheno$pd1 <- (HKU3$pheno$d1/HKU3$pheno$d0)*100
HKU3$pheno$pd2 <- (HKU3$pheno$d2/HKU3$pheno$d0)*100
HKU3$pheno$pd3 <- (HKU3$pheno$d3/HKU3$pheno$d0)*100
HKU3$pheno$pd4 <- (HKU3$pheno$d4/HKU3$pheno$d0)*100
```

**Area under the curve**

Use `MESS::auc` function which calculates the AUC for the weight trajectory, then subtract that from 400 (total area beneath 100% line from 0-4). 

Any weight above 100% would increase the AUC and decrease the area-above-the-curve (so some AACs will be negative).

Metric calculation:

```{r}
HKU3 <- calc_auc(cross = HKU3, steps = c(0,1,2,3,4), col.name = "weight_aac", 
                    phenos = c('pd0', 'pd1', 'pd2', 'pd3', 'pd4'))
```


## Exploratory data analysis  

How much weight do HKU3-infected mice lose on average? 

```{r}
pctloss <- (1-(HKU3$pheno$d4/HKU3$pheno$d0))*100
mean(pctloss)
range(pctloss)
```

Plot weight trajectory: 

```{r}
wtloss <- cov_trajectory_plot(HKU3, phenos = c('pd0', 'pd1', 'pd2', 'pd3', 'pd4'), 
                              ylab = "% of starting weight", ylim = c(72,114), 
                              title = "HKU3-CoV MA",parent.lty=2)
wtloss
```

Save plot:

```{r}
ensure_directory("figures/HKU3")
png(filename = "figures/HKU3/wt_loss.png", width = 700)
wtloss + bw_big_theme
dev.off()
```

Save for combining with other figures:

```{r}
wtloss <- wtloss + theme(legend.position = "none") + bw_big_theme 
ensure_directory("derived_data/otherRobjects")
saveRDS(wtloss, file = "derived_data/otherRobjects/hku3wtloss.Rdata")
```


### Hemorrhage score

```{r}
plotPheno(HKU3, pheno.col = 'HS')
```

Average HS:

```{r}
mean(HKU3$pheno$HS)
range(HKU3$pheno$HS)
median(HKU3$pheno$HS)
```

Data are normal enough, not transforming 

## Covariate analysis 

```{r}
anova(aov(d1 ~ d0 + batch + sex, data = HKU3$pheno))
anova(aov(d2 ~ d0 + batch + sex, data = HKU3$pheno))
anova(aov(d3 ~ d0 + batch + sex, data = HKU3$pheno))
anova(aov(d4 ~ d0 + batch + sex, data = HKU3$pheno))
anova(aov(weight_aac ~ batch + sex, data = HKU3$pheno))
anova(aov(HS ~ batch + sex, data = HKU3$pheno))
```

Baseline weight and sex, but not batch, affect outcome. 

## Define covariates 

Covariates: 
* Sex 
* Cross direction? Use parent_info.xlsx 

```{r}
HKU3$pheno$batch10 <- as.numeric(pull.pheno(HKU3, "batch") == 10)
HKU3$pheno$batch13 <- as.numeric(pull.pheno(HKU3, "batch") == 13)

# covar <- cbind(as.numeric(pull.pheno(HKU3, "sex") == "M"),
#                as.numeric(pull.pheno(HKU3, "batch") == 10), # batch 8 is baseline
#                as.numeric(pull.pheno(HKU3, "batch") == 13),
#                pull.pheno(HKU3, 'd0'))
# colnames(covar) <- c('sex', 'batch10', 'batch13', 'd0')
```


Get p-value for association with chr9:

* d2 ~ d0 + sex + chr9 
* d3 ~ d0 + sex + chr9
* d4 ~ d0 + sex + chr9
* aac ~ sex + chr9
* hs ~ sex + chr9 

These are the peak chr9 markers:

* S3C094905310 (SARS1: hs)
* SSH094942479 (SARS2: hs)
* UNC16977752 (SARS1: d2)
* S3N094839317 (SARS1: d3, aac; SARS2: d2, aac) <- use this one (most associations)
* SSH094942479 (SARS1: d4)
* gUNC17242574 (SARS2: d3)

```{r}
HKU3 <- fill.geno(HKU3)
HrS43m <- 'S3N094839317'
HrS43 <- pull.geno(HKU3, 9)[,HrS43m]
```

Day 2 weight: 

```{r}
lmod2 <- lm(d2 ~ d0 + sex + batch10 + batch13 + HrS43, HKU3$pheno)
summary(lmod2)
```

Day 3 weight:

```{r}
lmod3 <- lm(d3 ~ d0 + sex + batch10 + batch13 + HrS43, HKU3$pheno)
summary(lmod3)
```

Day 4 weight:

```{r}
lmod4 <- lm(d4 ~ d0 + sex + batch10 + batch13 + HrS43, HKU3$pheno)
summary(lmod4)
```

Weight AAC:

```{r}
lmodaac <- lm(weight_aac ~ sex + batch10 + batch13 + HrS43, HKU3$pheno)
summary(lmodaac)
```

HS:

```{r}
lmodhs <- lm(HS ~ sex + batch10 + batch13 + HrS43, HKU3$pheno)
summary(lmodhs)
```


## Phenotype-by-genotype plots 

Weight AAC vs chr9 

```{r}
p <- pxg(cross = HKU3, pheno = HKU3$pheno$weight_aac, marker = HrS43m,
     geno.map = list(A = "CC006", B = "CC044"), theme = sbs_pxg_theme, 
     qtl.map = data.frame(chr = c(9), qtl_name = c('HrS43')),
     title = "HKU3-CoV MA", ylim = c(-20,60))
```

Save jpeg 

```{r}
ensure_directory("figures/Fig4")
jpeg(filename = "figures/Fig4/hku3-HrS43-aac.png", width = 600)
p
dev.off()
```


Make Figure 4:

Load PxG from SARS1 and SARS2:

```{r}
pxg1 <- readRDS(file = "derived_data/otherRobjects/sars1-pxg-HrS43-aac.Rdata")
pxg2 <- readRDS(file = "derived_data/otherRobjects/sars2-pxg-HrS43-aac.Rdata")
```

```{r}
png("figures/Fig4/Figure4.png", height = 400, width = 1300)
plot_grid(pxg1, pxg2, p, align = "h", rel_widths = c(0.35,0.325,0.325),
          ncol = 3, labels = c('A', 'B', 'C'), label_size = 24)
dev.off()
```


What QTL from SARS1/SARS2 (other than *HrS43*) are associated with outcome in HKU3-infected mice? 

* *HrS44*: chr7, marker SBT072729469 (SARS-CoV-2) - associated with weight loss (d3: P=0.02, d4: P=0.05, AAC: P=0.02)
* *HrS45*: chr12, marker gUNC20800862 (SARS-CoV-2) - not associated
* *HrS46*: chr19, markers SFT190267185 (SARS-CoV) - not associated
* *HrS47*: chr15, marker gUNC26048180 (SARS-CoV-2) - not associated 
* *HrS48*: chr3, marker gUNC6002742 (SARS-CoV) - not associated
* *HrS49*: chr2, marker SFT021381098 (SARS-CoV) - associated with HS (P=0.03)


*HrS44* chr7, marker SBT072729469

```{r}
HrS44m <- 'SBT072729469'
HrS44 <- pull.geno(HKU3, 7)[,HrS44m]
```

Day 2 weight: 

```{r}
lmod2 <- lm(d2 ~ d0 + sex + batch10 + batch13 + HrS44, HKU3$pheno)
summary(lmod2)
```

Day 3 weight:

```{r}
lmod3 <- lm(d3 ~ d0 + sex + batch10 + batch13 + HrS44, HKU3$pheno)
summary(lmod3)
```

Day 4 weight:

```{r}
lmod4 <- lm(d4 ~ d0 + sex + batch10 + batch13 + HrS44, HKU3$pheno)
summary(lmod4)
```

Weight AAC:

```{r}
lmodaac <- lm(weight_aac ~ sex + batch10 + batch13 + HrS44, HKU3$pheno)
summary(lmodaac)
```

HS:

```{r}
lmodhs <- lm(HS ~ sex + batch10 + batch13 + HrS44, HKU3$pheno)
summary(lmodhs)
```


*HrS45* chr12, marker gUNC20800862

```{r}
HrS45m <- 'gUNC20800862'
HrS45 <- pull.geno(HKU3, 12)[,HrS45m]
```

Day 2 weight: 

```{r}
lmod2 <- lm(d2 ~ d0 + sex + batch10 + batch13 + HrS45, HKU3$pheno)
summary(lmod2)
```

Day 3 weight:

```{r}
lmod3 <- lm(d3 ~ d0 + sex + batch10 + batch13 + HrS45, HKU3$pheno)
summary(lmod3)
```

Day 4 weight:

```{r}
lmod4 <- lm(d4 ~ d0 + sex + batch10 + batch13 + HrS45, HKU3$pheno)
summary(lmod4)
```

Weight AAC:

```{r}
lmodaac <- lm(weight_aac ~ sex + batch10 + batch13 + HrS45, HKU3$pheno)
summary(lmodaac)
```

HS:

```{r}
lmodhs <- lm(HS ~ sex + batch10 + batch13 + HrS45, HKU3$pheno)
summary(lmodhs)
```


*HrS46* chr19, markers SFT190267185

```{r}
HrS46m <- 'SFT190267185'
HrS46 <- pull.geno(HKU3, 19)[,HrS46m]
```

Day 2 weight: 

```{r}
lmod2 <- lm(d2 ~ d0 + sex + batch10 + batch13 + HrS46, HKU3$pheno)
summary(lmod2)
```

Day 3 weight:

```{r}
lmod3 <- lm(d3 ~ d0 + sex + batch10 + batch13 + HrS46, HKU3$pheno)
summary(lmod3)
```

Day 4 weight:

```{r}
lmod4 <- lm(d4 ~ d0 + sex + batch10 + batch13 + HrS46, HKU3$pheno)
summary(lmod4)
```

Weight AAC:

```{r}
lmodaac <- lm(weight_aac ~ sex + batch10 + batch13 + HrS46, HKU3$pheno)
summary(lmodaac)
```

HS:

```{r}
lmodhs <- lm(HS ~ sex + batch10 + batch13 + HrS46, HKU3$pheno)
summary(lmodhs)
```


*HrS47* chr15 gUNC26048180

```{r}
HrS47m <- 'gUNC26048180'
HrS47 <- pull.geno(HKU3, 15)[,HrS47m]
```

Day 2 weight: 

```{r}
lmod2 <- lm(d2 ~ d0 + sex + batch10 + batch13 + HrS47, HKU3$pheno)
summary(lmod2)
```

Day 3 weight:

```{r}
lmod3 <- lm(d3 ~ d0 + sex + batch10 + batch13 + HrS47, HKU3$pheno)
summary(lmod3)
```

Day 4 weight:

```{r}
lmod4 <- lm(d4 ~ d0 + sex + batch10 + batch13 + HrS47, HKU3$pheno)
summary(lmod4)
```

Weight AAC:

```{r}
lmodaac <- lm(weight_aac ~ sex + batch10 + batch13 + HrS47, HKU3$pheno)
summary(lmodaac)
```

HS:

```{r}
lmodhs <- lm(HS ~ sex + batch10 + batch13 + HrS47, HKU3$pheno)
summary(lmodhs)
```


*HrS48* chr3, marker gUNC6002742

```{r}
HrS48m <- 'gUNC6002742'
HrS48 <- pull.geno(HKU3, 3)[,HrS48m]
```

Day 2 weight: 

```{r}
lmod2 <- lm(d2 ~ d0 + sex + batch10 + batch13 + HrS48, HKU3$pheno)
summary(lmod2)
```

Day 3 weight:

```{r}
lmod3 <- lm(d3 ~ d0 + sex + batch10 + batch13 + HrS48, HKU3$pheno)
summary(lmod3)
```

Day 4 weight:

```{r}
lmod4 <- lm(d4 ~ d0 + sex + batch10 + batch13 + HrS48, HKU3$pheno)
summary(lmod4)
```

Weight AAC:

```{r}
lmodaac <- lm(weight_aac ~ sex + batch10 + batch13 + HrS48, HKU3$pheno)
summary(lmodaac)
```

HS:

```{r}
lmodhs <- lm(HS ~ sex + batch10 + batch13 + HrS48, HKU3$pheno)
summary(lmodhs)
```


*HrS49* chr2, marker SFT021381098

```{r}
HrS49m <- 'SFT021381098'
HrS49 <- pull.geno(HKU3, 2)[,HrS49m]
```

Day 2 weight: 

```{r}
lmod2 <- lm(d2 ~ d0 + sex + batch10 + batch13 + HrS49, HKU3$pheno)
summary(lmod2)
```

Day 3 weight:

```{r}
lmod3 <- lm(d3 ~ d0 + sex + batch10 + batch13 + HrS49, HKU3$pheno)
summary(lmod3)
```

Day 4 weight:

```{r}
lmod4 <- lm(d4 ~ d0 + sex + batch10 + batch13 + HrS49, HKU3$pheno)
summary(lmod4)
```

Weight AAC:

```{r}
lmodaac <- lm(weight_aac ~ sex + batch10 + batch13 + HrS49, HKU3$pheno)
summary(lmodaac)
```

HS:

```{r}
lmodhs <- lm(HS ~ sex + batch10 + batch13 + HrS49, HKU3$pheno)
summary(lmodhs)
```
